### 2015-03-23 resistanceModelNotesAndy.Rmd
noting progress

---
title: "resistanceModelNotesAndy"
output: html_document
---

Notes on insecticide resistance model for liverpool.
andy south 27/1/15

My first go at keeping notes like this as an rmarkdown file.

**27/1/15**  7.5hrs

```
#working version of Beths code
#run it outside of this project because it changes wd
source("C:\\Dropbox\\Ian and Andy\\andy\\malaria\\Beth code\\malaria_code_beth.r")
```

Editing `malaria_code_andy.r`, indenting functions etc.

Much of the script from line 464 is setting up parameters for example runs.

input[a,b] : a=parameter number and b=scenario number

464         set input params
1058-2716   run model
2718        Actions needing full results.list (e.g. curtis plots use multiple scenarios)

edited bookmarks to navigate around the file in RStudio.

**9/3/15** 4-5 1hr

**10/3/15** 4.15 - 7.15 3hrs

```{r}
listFunctions <- function(filename) {
  temp.env <- new.env()
  sys.source(filename, envir = temp.env)
  functions <- lsf.str(envir=temp.env)
  rm(temp.env)
  return(functions)
}

#listFunctions("C:\\Dropbox\\Ian and Andy\\andy\\malaria\\Beth code\\malaria_code_beth.r")
```

So there are only 11 functions.

allele.freq : function (mat)  
curtis_f1 : function (nrelaxmat, relaxmat, gencol, r1col)  
curtis_f2 : function (combmat, bmat, amat, gencol, r1col, r2col)  
curtis_ld : function (resultsmat, relaxedmat, gencol, ldcol)  
haplotype : function (mat)  
HW : function (P, mat)  
linkage : function (mat)  
make.genotypemat : function (P_1, P_2)  
make.matrix : function (mat, rnames)  
singlealleleFrequency : function (locus, max_gen, results.list, input)  
timetoFifty : function (locus, max_gen, results.list, input

I can create an initial github repo for the files from Beth (&this notes file)
That means I will have version control for my initial restructuring. Later I may want to create a new repo for the reformatted code as a package.

What to call the first repo ? resistance

1. created repo on github
1. created RStudio project from github repo
+ To get ssh push working.
+ RStudio Tools, Shell
+ git remote set-url origin git@github.com:AndySouth/resistance.git

**11/3/15** 9.30-12.30 3hrs 2.15-6.45 7.5hrs

moved my dropbox folder back to C.

Plan
1. move Beths functions out into a single file for each in an R folder
1. edit Beths commments into roxygen format
1. only later think about changing functions and documentation to follow best practice

Changed plan slightly by appending plot onto start of plot functions.

To source files in the R folder
`lapply(dir("R"),function(x) source(paste0("R//",x)))`

**12/3/15** 9.30-14 4.5hrs 3-5 6-7 7.5hrs

The model generates 3 matrices : results, genotype, fitness
1. Results : freq of R allele at each loci in each sex and linkage disequilibrium of R allele in each sex, per generation
2. Genotype : frequencies of each of the ten genotypes, per generation
3. Fitness : fitness scores of each genotype/niche combination (table 4. of Main Document)

When a file is input with multiple scenarios, each of the three matrices is stored in a list, where scenario number gives position of the matrix in the list.

results.list : 
fitness.list : 
genotype.list :

Can I put these into a single list so that they can be returned from a single function ?

done ~ put input object creation into a function
But there is still a fair bit of code that reads parameter values out of the input matrix and into named variables.

I might be able to restructure these collections of single variables into arrays.
e.g. for the exposure levels of m&f

```{r}
#to create an array a[sex][locus1][locus2]
sex <- c("F","M")
locus1 <- c("0","A","a")
locus2 <- c("0","B","b")
dimnames1 <- list( sex=sex, locus1=locus1, locus2=locus2 )
dim1 <- sapply(dimnames1, function(x) length(x))
a <- array(0,dim=dim1, dimnames=dimnames1)
a
length(a)
#to access one element
a['M','a','b']
#to check that male exposures total 1
sum(a['M',,])
```

```
  ## Exposure levels of males and females to each insecticide niche ##
  # males
  a.m_00 <- input[8,i]

  a.m_a0 <- input[9,i]
  a.m_A0 <- input[10,i]

  a.m_0b <- input[11,i]
  a.m_0B <- input[12,i]

  a.m_ab <- input[13,i]
  a.m_AB <- input[14,i]

  a.m_Ab <- input[15,i]
  a.m_aB <- input[16,i]

  #a.m <- sum(a.m_00, a.m_a0, a.m_A0, a.m_0b, a.m_0B, a.m_ab, a.m_AB, a.m_Ab, a.m_aB)
  #if ( a.m != 1 ){		 
  #	print( paste("Error in male exposures: must total one: ", a.m) )

  #	}else{
  #		print( paste( "Male exposures total 1: ", a.m ))
  #		}

  # females
  a.f_00 <- input[17,i]

  a.f_a0 <- input[18,i]
  a.f_A0 <- input[19,i]

  a.f_0b <- input[20,i]
  a.f_0B <- input[21,i]

  a.f_ab <- input[22,i]
  a.f_AB <- input[23,i]

  a.f_Ab <- input[24,i]
  a.f_aB <- input[25,i]
```

I should check that what we are doing fits in with what others are doing. We may be able to capitalise on existing tools.

[popgen on CRAN](http://cran.r-project.org/web/packages/popgen/popgen.pdf)
seems a bit old and doesn't do a huge amount, no vignette.

[gstudio](http://cran.r-project.org/web/packages/gstudio/) looks more promising. e.g. it defines locus objects.
[gstudio documentation](http://dyerlab.github.io/gstudio/)
A locus object is set like this, but I'm not sure if this would help us.
It does have some stuff on drift at end of help.
```
require(gstudio)
loc <- locus(c("C", "A"))
loc
```

pegas provides functions for the analysis of allelic data and of haplotype data from DNA sequences.
It requires and complements two other R-packages: ape and adegenet.
[pegas](http://ape-package.ird.fr/pegas.html)
[pegas data structures](http://ape-package.ird.fr/pegas/DefinitionDataClassesPegas.pdf)
includes a class called loci. : An object of class "loci" is a data frame where rows represent individuals
and columns are loci and optional additional variables.
Again I can't quite see how it would be useful for us.

There is a [HardyWeinberg](http://cran.r-project.org/web/packages/HardyWeinberg/vignettes/HardyWeinberg.pdf) package on CRAN. "The HardyWeinberg package consists of a set of tools for analyzing diallelic
genetic markers, and is particularly focused on the graphical representation of their (dis)equilibrium condition in various ways."

There is a CRAN [genetics task view](http://cran.r-project.org/web/views/Genetics.html).

The [genetics package](http://cran.r-project.org/web/packages/genetics/genetics.pdf) which also has a locus class, seems to be more detailed, about chromosomes etc. and probably not of use to us.

#### Structure of the results matrices

```{r}
max_gen <- 2 #just as example

# Set up results matrix - prints overall freq of R and S allele per locus per sex, LD and overall allele freq (i.e. 1)
results <- matrix ( nrow = max_gen, ncol = 11 )
colnames( results ) <- c( "Gen", "m.R1", "m.R2", "m.LD","f.R1", "f.R2", "f.LD", "M", "F", "dprime", "r2" )

# set up fitness by niche matrix - records fitness scores for each niche for each genotype
fitness <- matrix ( nrow = 10, ncol = 9, c(rep(0,90)))
colnames(fitness) <- c( "-,-", "a,-", "A,-", "b,-", "B,-", "a,b", "A,B", "A,b", "a,B" )
rownames(fitness) <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
						"RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
						"RR1SS2", "RR1RS2", "RR1RR2")

# set up genotype matrix - records frequencies of each of the 9 two locus genotypes each generation
genotype <- matrix( nrow=max_gen, ncol=11 )
colnames(genotype) <- c("gen", "SS1SS2", "SS2RS2", "SS1RR2", 
						"RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
						"RR1SS2", "RR1RS2", "RR1RR2")

```

Somehow it seems I've broken it, I'm getting NAs in results matrix.

The input files are different (but same structure) between Beths version that works and my new version ...
calibration is 100 in my failing version and 1012 in Beths, i think this is because I'm reading in the csv when I shouldn't.

I seem not to get to line 889 in my version
  	# male
  	f.m.SS1SS2 <- genotype.freq[1,]

Then suddenly it seemed to start working when I put a browser in.
It's something tricky around line 711, if it doesn't get to 889 it doesn't fill in the results matrix ??

Aha! It seems to have been this without the comments before the tilde's that stoppped it from working
```
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 		
  ### Loop to run the model from the initial conditions generated above ####
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
```

**13/3/15** 9.15-12.15 3hrs

A reproducible version of yesterdays bug.

```
x<-1
~
for(i in 1:10) x<-x+1
cat(x)
```
[R bug reporting](https://bugs.r-project.org/bugzilla3/enter_bug.cgi)

1. can I create a list of the 3 lists for results ?
Yes
Replace this  
results.list <- list()			
fitness.list <- list()
genotype.list <- list()	
with  
listOut <- list( results=list(), fitness=list(), genotype=list() )

Then find & replace  
results.list with listOut$results  
fitness.list with listOut$fitness
genotype.list with listOut$genotype

**16/3/15** mon 18.30-19.30 1hr
cut out lines 92 to 1740 into a runModel() function, later can subdivide it up

**17/3/15** tues to liverpool 8.30-5.30 9hrs

Assuming that resistance to each insecticide was encoded at one locus respectively, and that these can either have a homozygous susceptible (SS), heterozygous (RS) or homozygous resistant (RR) genotypes at each locus, there are then ten possible genotypes (including cis & trans forms of the double heterozygous).

We consider two insecticides (A and B) at three possible levels; absent, low(ab) and high(AB), resulting in 9 niches;

Locus1 relates to resistance for insecticide A, and locus2 for B. Locus2 does not effect fitness under exposure to A.

Males and females can be exposed differently.

Curtis states that if resistance to one insecticide is present at very low levels in the population, then resistance will rise slower if this insecticide is used on its own, rather than in a combination with a more established insecticide.

This is the important Curtis statement “the use of a mixture where the initial gene frequencies are unequal leads to more rapid increase in the frequency of the rarer of the genes”

A key aspect of Curtis’ prediction was that this relationship would be determined by an increasing association between the presence of the two resistance alleles, measured as high levels of linkage disequilibrium.

The results given above seem to be in contrast with Curtis’ assumption, and suggest that in fact using the insecticides in combination can slow the spread of resistance.

My summary :
Curtis suggests resistance spreads more quickly with 2 insecticides, due to  increasing association between the alleles. Beths work suggests that combination can slow spread of resistance.

Good figure 1 in manuscript :
The rate of spread of IR depends on whether the resistance mutation is dominant, semi-dominant, or recessive.

**aha! I've only just understood this bit**
**At high concentrations only the RR individuals survive so resistance is recessive; as concentrations decline some RS mosquitoes survive making resistance semi-dominant; at low concentration both RR and RS mosquitoes survive, making resistance dominant.** 
**so main problems of resistance arise when insecticide concentrations decline**

From manuscript about scenarios to run:
**row in params csv**

Insecticide coverage for females: 0.1, 0.2, 0.6, 0.8, 0.9 (n=5) **9 & 15**  
Insecticide coverage for males: as for females, or half that of females (n=2) **18 & 24**  
‘New’ insecticide resistance starting freq: 0.0001, 0.001, 0.01 (n=3) **6**  
‘Old’ insecticide resistance starting freq: 0.05, 0.1, 0.2, 0.5, 0.8 (n=5) **7**  
Resistance heterozygous fitness: 0.5, 0.75. 1    (n=3 for each insecticide) **?28 & 30**  
Sensitive heterozygous fitness:0, 0.2, 0.4 (n=3 for each insecticide) **?27 & 29**  
Resistance allele dominance: 0, 0.2, 0.5, 0.9, 1 for each locus (n=5 for each locus) **35,38 do to start**  
Fitness costs: 0, 0.05, 0.1 (n=3) **43&44**  
Fitness dominance: 0, 0.5, 1 (n=3) **33&36**

BUT now we are going to do it by sampling from a uniform distribution before doing decision trees. Then you can progressively add results to a file.

I could start by creating a document with the results at the extremes of the parameter space.

Send Ian & Beth devtools installation instructions and a document.

Approx. 60,000 combinations. Too many???

would latin cube hypersampling be better?

The simulation output was the ratio of  the time (in generation) for resistance  to the ‘new’ insecticide  to reach a given allele threshold frequency when the ‘new’ insecticide was deployed as combination, compared to the time taken to reach the same threshold if deployed on its own; a ratio >1 indicates that combination is better. Three threshold frequencies were investigated: 0.1, 0.25, 0.5.

These ratios was then analysed by regression to find out how the input parameters alter the ratio, and hence under what circumstance combination would be best/worst. Classification tress can also be done.

The intention was not to consider particular plausible deployment situations, but was to systematically investigate a wide range of parameter space to determine whether CD was better than single-insecticide deployment at reducing the rate at which the minor resistance form spreads.

Show my progress :

1. github repo
1. move functions to their own files and structure documentation
1. move parts of main file into functions
1. combine 3 outputs to a single list so that can be returned from a function
1. nearly able to create a package with structured documentation etc.

Questions for Beth & Ian

1. Curtis Table1 vi that shows resistance spreading faster with mixture.
1. is coverage same as exposure ? yes
1. what subset of parameters would they like me to vary to start
1. locations of sensitivity params in params.csv
1. what outputs are wanted for sensitivity scenarios ? num generations until resistance reaches threshold (0.1, 0.25 & 0.5) Curtis just used 0.5.

Talking with Ian & Beth

Sequential use.

For some scenarios, we'll need to run with one insecticide and then stop and start other.

When does mixture fail, until one resistance allele reaches critical point, then continue using 2nd insecticide until resistance for that one reaches critical point too.

Could create a version of timetoFifty that accepts threshold param.

Ian has created list of defaults in most recent manuscript.

The Curtis version is going to be simpler. Because not all param variations are included.

**17/3/15** weds in liverpool 8.30-6 9.5hrs

Ian created `suggested_sensitivity_analysis.doc`

I created a variables table in there with Ians parameter ranges.

My poor paraphrasing of what I think Ian said about the sequential scenarios. The starting conditions from after one insecticide has 'failed' are the same as starting from scratch for the other insecticide because the insecticides to do effect selection on the other locus.

moving `malaria_code_andy.r` into a function resistanceMaster()
exposed a couple of variable scoping problems.

plus now the graphs look different and I get :
Warning message: In data.matrix(mat) : NAs introduced by coercion

this is from createInputMatrix()
input <- make.matrix(input, input$Input)

this might be something to do with the location of "input.parameters.csv" which I need to sort ...

devtools commands for setting up a package :
create requires that the directory doesn't exist yet; it will be created. 
setup assumes an existing directory from which it will infer the package name.

probably need to set rstudio rg to false because mine is already a rs project.

setup(rstudio=FALSE)

Then in RStudio I had to set project options, build tools to package
(so maybe I should have done just done detools::setup())

but on build :
ERROR: The build directory does not contain a DESCRIPTION
file so cannot be built as a package.
Build directory: C:/rsprojects/resistance

but there is a DESCRIPTION there ???

devtools::load_all(".")
Error: Line starting 'person("Bethany", "L ...' is malformed!
sorted, just needed tab before 2ry authors

package does build now

potential refactoring of code, if i think carefully, using arrays I should be able to replace this :
```
    if(calibration==103){		## no selection calibration
        ## male
        # SS1
        fs.m.SS1SS2 <- f.m.SS1SS2
        fs.m.SS1RS2 <- f.m.SS1RS2
        fs.m.SS1RR2 <- f.m.SS1RR2
        # RS1
        fs.m.RS1SS2 <- f.m.RS1SS2 
        fs.m.RS1RS2_cis <- f.m.RS1RS2_cis
        fs.m.RS1RS2_trans <- f.m.RS1RS2_trans
        fs.m.RS1RR2 <- f.m.RS1RR2
        # RR2 
        fs.m.RR1SS2 <- f.m.RR1SS2
        fs.m.RR1RS2 <- f.m.RR1RS2
        fs.m.RR1RR2 <- f.m.RR1RR2

        ## female
        # SS1
        fs.f.SS1SS2 <- f.f.SS1SS2
        fs.f.SS1RS2 <- f.f.SS1RS2
        fs.f.SS1RR2 <- f.f.SS1RR2
        # RS1
        fs.f.RS1SS2 <- f.f.RS1SS2 
        fs.f.RS1RS2_cis <- f.f.RS1RS2_cis
        fs.f.RS1RS2_trans <- f.f.RS1RS2_trans
        fs.f.RS1RR2 <- f.f.RS1RR2
        # RR2 
        fs.f.RR1SS2 <- f.f.RR1SS2
        fs.f.RR1RS2 <- f.f.RR1RS2
        fs.f.RR1RR2 <- f.f.RR1RR2
```

with something like :
`fs <- f`

where arrays are set up as :
```{r}
#to create an array a[sex][locus1][locus2]
namesLoci <- c('SS','RS','RR')
locus1 <- paste0(namesLoci,'1')
locus2 <- paste0(namesLoci,'2')
sex <- c("F","M")

dimnames1 <- list( sex=sex, locus1=locus1, locus2=locus2 )
dim1 <- sapply(dimnames1, function(x) length(x))
a <- array(0,dim=dim1, dimnames=dimnames1)
a
length(a)
#to access one element
a['M','SS1','RR2']
#to check that males total 1
sum(a['M',,])
```

Just need to work out how best to deal with cis & trans.
This would be a less pleasing alternative
```{r}

namesLoci <- c('SS1SS2','SS1RS2','SS1RR2',
               'RS1SS2','RS1RS2','RS1RR2',
               'RR1SS2','RR1RS2','RR1RR2')

#adding cis & trans
namesLoci <- c('SS1SS2','SS1RS2','SS1RR2',
               'RS1SS2','RS1RS2cis','RS1RS2trans','RS1RR2',
               'RR1SS2','RR1RS2','RR1RR2')

sex <- c("F","M")

dimnames1 <- list( sex=sex, namesLoci=namesLoci )

dim1 <- sapply(dimnames1, function(x) length(x))
a <- array(0,dim=dim1, dimnames=dimnames1)
a
length(a)
#to access one element
a['M','SS1SS2']
#to check that males total 1
sum(a['M',])
str(a)
```

**19/3/15** fri 9.45-14 4.25hrs 3.30-7 3.5hrs 7.75hrs

checking whether I can run without the input file. Fixed now.

```{r}
resistanceMaster(params.csv = FALSE)
```

This is what the fitness scores that can be saved for each scenario as *two-locus_fitness-scores.csv look like :
```
	-,-	a,-	A,-	-,b	-,B	a,b	A,B	A,b	a,B
SS1SS2	1	0	0	0	0	0	0	0	0
SS1RS2	1	0	0	0	0	0	1.86E-05	0	0
SS1RR2	1	0	0	0	0	0	0.1161	0	0
RS1SS2	1	0	0	0	0	0	0	0	0
RS1RS2	1	0	0	0	0	0	2.13E-05	0	0
RS1RR2	1	0	0	0	0	0	0.132913	0	0
RR1SS2	1	0	0	0	0	0	0	0	0
RR1RS2	1	0	0	0	0	0	3.44E-05	0	0
RR1RR2	1	0	0	0	0	0	0.215	0	0
```

Starting resistSimple() to be able to run single simple Scenarios.

**20/3/15** sat 10.15-2.15 4hrs

[recent popgen hack](https://github.com/NESCent/r-popgen-hackathon/wiki)
Doesn't seem that useful for this project.

Good progress :
check out e.g.
`resistSimple(P_1=0.01,P_2=0.3)`

**22/3/15** mon 9.45

### TO DO

sort this
To git@github.com:AndySouth/resistance.git
 ! [rejected]        master -> master (non-fast-forward)
error: failed to push some refs to 'git@github.com:AndySouth/resistance.git'
hint: Updates were rejected because the tip of your current branch is behind
hint: its remote counterpart. Integrate the remote changes (e.g.
hint: 'git pull ...') before pushing again.
hint: See the 'Note about fast-forwards' in 'git push --help' for details.

1. create setInputOneScenario to enable setting inputs for a single scenario
+ sensible defaults
+ later will allow customised scenarios e.g. setInputCurtisFig1
+ will reduce code volume and repetition in createInputMatrix

1. finish tidying documentation of reading in input vars in createInputMatrix() add avr names from runModel()
+ can use these later for roxygen doc of function inputs to resistSimple
+ and I could use find&replace to rename later if needed

1. check whether setting input[13,1] <- to the next line at 58 in createInputMatrix() was a bug

1. sort that resistSimple will only work with calibration of 1012

1. make it easier to run a single scenario and default scenarios
+ might do this by first adding a function rather than changing existing stuff
1. i think the main visual output is the frequency of resistance over time
1. start creating a vignette to help me understand how the model works

1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
1. put back in checks for things that should sum to 1
1. edit runModel carefully
1. check out the part about running sequential insecticide scenarios
1. start to write script to run sensitivity analysis, does it want to work by creating a large input.csv file, or by setting input object directly.
1. spend some time developing my understanding of how the model works
1. submit tilde bug report
1. later, structure as R package, maybe move to a new github repo
1. re-look at R popgen packages when I've got more of a handle on what we are doing.

### suggestion from Beth
a solution may be to provide a .csv file with the Curtis scenarios set up
and then the need to set the calibration number and params.csv in the R script can be removed altogether
the only bit of R code the user will need to change is the setwd() and the read.csv()

### variable naming conventions
**SS1, RS1, RR1** homozygous susceptible, heterozygous or homozygous resistant genotypes at each locus.   
With 2 loci there are ten possible genotypes (including cis & trans forms of the double heterozygous).

** A, B, a, b, -** 2 insecticides at 3 levels; low(ab), high(AB), absent(--) resulting in 9 niches.

Locus1 relates to resistance for insecticide A, and locus2 for B. Locus2 does not effect exposure to A.

```
colnames(fitness) <- c( "-,-", "a,-", "A,-", "b,-", "B,-", "a,b", "A,B", "A,b", "a,B" )
rownames(fitness) <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
                        "RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
                        "RR1SS2", "RR1RS2", "RR1RR2")
```