### 2016-05-16 paper2_resistance_mechanisms_mixtures.Rmd
updating paper2

---
title: "paper2_resistance_mechanisms_mixtures"
output: pdf_document
#output: word_document
author: "Andy South, Ian Hastings and Bethany Levick"
date: "`r Sys.Date()`"
---

```{r,echo=FALSE,results='hide'}
  library(resistance)
```

### version 1

# Mechanisms by which insecticide resistance can be influenced by strategies of insectcide use : a simple theoretical exploration.

#### (This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
#### Target : Malaria journal

## Abstract

This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence/rotation and mixtures. It is aimed at a readership with limited experience of population genetics and resistance. It uses a model representing two independent loci coding for resistance to two insecticides. It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective adavantage of resistance and the dominance of the alleles. There is uncertainty about the values of these parameters in field situations. We demonstrate how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies. We describe an R package and user-interface that allows further exploration of the development of resistance.

## Introduction

Insecticide resistance is a problem for malaria (GPIRM), other vector borne diseases and agriculture.

Insecticide resistance becomes a problem when genes coding for resistance firstly arise in a population and secondly increase in frequency. We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.

The changing frequency of insecticide resistance is a population genetic process that can be influenced by :

Parameter  | Description
------------------- | -------------------------------------------------------
1. Effectiveness | proportion of exposed insects with the susceptible allele that are killed by the insecticide
2. Exposure | proportion of the insect population that are exposed to the insecticide
3. Dominance | dominance of the resistance allele relative to the susceptible allele
4. Selection Coefficient  | selective advatange of resistance in the presence of the insecticide
5. Frequency  | frequency of resistance alleles within the population

We use a 2-locus population genetic model of insecticide resistance (Hastings et al. in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies. Hastings et al. (in press) describe the technical details of the model and explore large parameter ranges. In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate the mechanisms by which different parameter values are expected to influence the devlopment of resistance.

## Methods

*accessible summary of the model

## Results

*Maybe start by exploring the influence of the parameters on sole use *
*could put here or in the discussion

**single insecticide application effect on increase in resistance of increasing each parameter**

Parameter  | effect on resistance | Mechanism
------------------ | ------- | ---------------------------------------------------
1. Effectiveness | faster | increased deaths of susceptibles
2. Exposure | faster | increased deaths of susceptibles
3. Dominance | faster | increased survival of resistants
4. Selection Coefficient | faster | increased survival of resistants
5. Frequency  | faster | less change needed to reach resistance threshold, and rate of change in resistance tends to increase with frequency

*to help explain I could start from a scenario where the attributes of each insecticide are the same*
*and then explore changing each in turn*
*I could then work up to the curtis fig2 plot*

**When all parameters are equal resistance arises slower for sequential than mixture.**
**Effect of changing parameters on relative performance of mixture versus sequential strategy**

Parameter  | relative performance mix versus seq | Mechanism
------------------ | ------- | ---------------------------------------------------
A. All equal | seq > mix | resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence
1. Effectiveness | faster | ...
2. Exposure | faster | ...
3. Dominance | faster | ...
4. Selection Coefficient | ...
5. Frequency  | faster | ...

### Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
#### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

#### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

#### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

## Discussion

(@)  In much of parameter space resistance changes more slowly when insecticides are used in a mixture relative to sequential use. The mechanism explaining this is ...

(@) For a mixture of 2 insecticides with differing effectiveness. Resistance to the insecticide with a greater effectiveness increases faster. The less effective insecticide is 'protected' by the more effective. Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.

(@) When the effectiveness of 2 insecticides in a mixture is similar, the more effective insecticide provides little 'protection' to the less effective. In this case resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence :

```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( max_gen=500, P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.17 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.75 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.23 , correctMixDeployProp = 1 , addCombinedStrategy = FALSE )
```

## References
1. GPIRM
1. Curtis 
1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
1. Birget 2015

## Todo

1. re-read Mani 1985
1. re-read GPIRM
1. mention Curtis but don't depend on it
1. do I want y axis of resistance frequency plots to be on a log scale ?
1. can I develop explanatory diagrams to describe the input parameters ?
1. decide on base starting scenarios and how to modify params from that
1. mention generation time of 1 month, maybe even rescale axes to years ?
1. look at effect of sex linkage