### 2016-06-29 inst/documents/testfigs_simpler.Rmd
failed attempt at getting fig.subcap to work

---
title: "This is my minimal example"
output:
  pdf_document:
    fig_caption: yes
  graphics: yes
  keep_tex: yes  
  header-includes:
    -\usepackage{subfig}

#includes:
#  in_header: styles.tex
---

My working plot.

```{r, results = 'asis', echo=FALSE, fig.width=3, fig.height=3, fig.cap="Figure 1: A smiley face because I work."}
plot(-5:5, (-5:5)^2)
```

My failing plot.

```{r, results = 'asis', echo=FALSE, fig.show='hold', fig.width=3, fig.height=3, fig.cap="Figure 2: I do not work :(", fig.subcap = c("Neither do I", "or me!")}
plot(1:10)
plot(rnorm(10))
```