### 2016-10-10 Matt/git-instructions-for-matt.Rmd
instructions for Matt

---
title: "git-instructions-for-matt"
author: "Andy South"
date: "10 October 2016"
output: html_document
---

Instructions for creating a fork of the resistance repository, being able to add your own updates, and to receive updates from me.

https://github.com/AndySouth/resistance

1) click fork
will create : https://github.com/Mapowney/resistance

2) click clone to get the URL to use in git clone
#from new local folder 
git clone git@github.com:Mapowney/resistance.git

3) set up syncing
cd resistance
git remote add upstream git@github.com:AndySouth/resistance.git

#to view remote repos
git remote -v

4) SYNC FORK from upstream, to get updates from me
git fetch upstream
git checkout master
git merge upstream/master

5) COMMIT & PUSH local changes to the fork, to put your own updates on github
(this step can be done from RStudio UI)
git commit -a -m "message" 
git push

6)