### 2016-10-14 inst/documents/effectiveness_exposure_difference.Rmd
start doc on diff between effectiveness & exposure

---
title: "Why is time-to-resistance in mixtures influenced differently by effectiveness and exposure ?"
output: pdf_document
#output: word_document
author: "Andy South"
date: "`r Sys.Date()`"
---

```{r,echo=FALSE,results='hide'}
library(resistance)
library(ggplot2)
```

Increasing the exposure to one insecticide in a mixture has liitle influence 
on the time-to-resistance of the other. In contrast increasing the effectiveness 
of one insecticide in a mixture increases the time-to-resistance of the other.

Why is there this difference between the effects of effectiviness and exposure in mixtures ?

For single insecticide application increasing effectiveness or exposure have the same effect
of decreasing time-to-resistance by increasing selection pressure.

```{r, eval=TRUE, echo=FALSE}

#setting up the data

df_base_ <- as.data.frame(fitnessGenotype(exp1=0.5, exp2=0.5, plot=FALSE))[2,]
df_base <- df_base_/max(df_base_)

df_ef_ <- as.data.frame(fitnessGenotype(exp1=0.5, exp2=0.5, eff1=0.8,  plot=FALSE))[2,]
df_ef <- df_ef_/max(df_ef_)
#df_ef

df_ex_ <- as.data.frame(fitnessGenotype(exp1=0.8, exp2=0.5, plot=FALSE))[2,]
df_ex <- df_ex_/max(df_ex_)
#df_ex

#to allow me to plot them

#normalised ones
df_base <- data.frame(t(df_base),row.names=row.names(t(df_base)))
df_ef <- data.frame(t(df_ef),row.names=row.names(t(df_ef)))
df_ex <- data.frame(t(df_ex),row.names=row.names(t(df_ex)))
#not normalised
df_base_ <- data.frame(t(df_base_),row.names=row.names(t(df_base_)))
df_ef_ <- data.frame(t(df_ef_),row.names=row.names(t(df_ef_)))
df_ex_ <- data.frame(t(df_ex_),row.names=row.names(t(df_ex_)))

```

```{r, eval=TRUE, echo=FALSE}

plot_fit_rs(df_base, column='f', title='base')

plot_fit_rs(df_ef, column='f', title='effectiveness')

plot_fit_rs(df_ex, column='f', title='exposure')

```

Seems that in the case where effectiveness is increased the fitnesses are closer together so does that mean that there is less selection for resistance ?

What is happening with the 2nd insecticide ?

#Normalised
```{r, eval=TRUE, echo=FALSE}

calc_fit_rs(df_base,'f',round_places=3)
#R2-S2  R2-S2 0.066

calc_fit_rs(df_ef,'f',round_places=3)
#R2-S2  R2-S2 0.045

calc_fit_rs(df_ex,'f',round_places=3)
#R2-S2  R2-S2 0.073
```

So yes, increasing effectiveness I1 results in a decrease in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.045), this can explain the slower selection for resistance.

In contrast increasing exposure to I1 results in an increase in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.073), this can explain the faster selection for resistance.

Be a bit careful, does my normalising do anything weird ? I might be able to run calc_fit_rs() on the non normalised versions.
not normalised
calc_fit_rs(df_base_,'f')
      allele    meanfit
R1        R1 0.721
S1        S1 0.669
R2        R2 0.721
S2        S2 0.669
R1-S1  R1-S1 0.052
#R2-S2  R2-S2 0.052

calc_fit_rs(df_ef_,'f')
      allele    meanfit
R1        R1 0.666
S1        S1 0.583
R2        R2 0.641
S2        S2 0.608
R1-S1  R1-S1 0.083
#R2-S2  R2-S2 0.033

calc_fit_rs(df_ex_,'f')
      allele    meanfit
R1        R1 0.633
S1        S1 0.531
R2        R2 0.608
S2        S2 0.556
R1-S1  R1-S1 0.102
#R2-S2  R2-S2 0.052

So in the non normalised version the difference in fitness for S&R I2 stays the same when I1 exposure is increased but decreases when I1 effectiveness is increased.

In text form :
In a mixture, increasing the exposure to one insecticide has very little effect on the selection pressure for resistance to the other. In contrast increasing the effectiveness of one insecticide in a mixture decreases the slection pressure for resistance to the other.