### 2016-10-17 inst/documents/effectiveness_exposure_difference.Rmd
adding examples & finessing case

---
title: "Why is time-to-resistance in mixtures influenced differently by effectiveness and exposure ?"
output: pdf_document
#output: word_document
author: "Andy South"
date: "`r Sys.Date()`"
---

```{r,echo=FALSE,results='hide'}
library(resistance)
library(ggplot2)
```

Increasing the exposure to one insecticide in a mixture has little influence 
on the time-to-resistance of the other. In contrast, increasing the effectiveness 
of one insecticide in a mixture increases the time-to-resistance of the other.

Why is there this difference between the effects of effectiviness and exposure in mixtures ?

For single insecticide application increasing effectiveness or exposure have the same effect
of decreasing time-to-resistance by increasing selection pressure.

scaneario exploration :

1. base : all inputs set to 0.5
1. effectiveness : effectiveness set to 0.8
1. exposure : exposure set to 0.8

```{r, eval=TRUE, echo=FALSE}

#setting up the data

df_base_ <- as.data.frame(fitnessGenotype(exp1=0.5, exp2=0.5, plot=FALSE))[2,]
df_base <- df_base_/max(df_base_)

df_ef_ <- as.data.frame(fitnessGenotype(exp1=0.5, exp2=0.5, eff1=0.8,  plot=FALSE))[2,]
df_ef <- df_ef_/max(df_ef_)
#df_ef

df_ex_ <- as.data.frame(fitnessGenotype(exp1=0.8, exp2=0.5, plot=FALSE))[2,]
df_ex <- df_ex_/max(df_ex_)
#df_ex

#to allow me to plot them

#normalised ones
df_base <- data.frame(t(df_base),row.names=row.names(t(df_base)))
df_ef <- data.frame(t(df_ef),row.names=row.names(t(df_ef)))
df_ex <- data.frame(t(df_ex),row.names=row.names(t(df_ex)))
#not normalised
df_base_ <- data.frame(t(df_base_),row.names=row.names(t(df_base_)))
df_ef_ <- data.frame(t(df_ef_),row.names=row.names(t(df_ef_)))
df_ex_ <- data.frame(t(df_ex_),row.names=row.names(t(df_ex_)))

```

# before normalisation
```{r, eval=TRUE, echo=FALSE, results='hide', fig.width=2}

plot_fit_rs(df_base_, column='f', title='base', ylim=c(0.4,0.8))

plot_fit_rs(df_ef_, column='f', title='effectiveness', ylim=c(0.4,0.8), yblank=TRUE)

plot_fit_rs(df_ex_, column='f', title='exposure', ylim=c(0.4,0.8), yblank=TRUE)

```

# normalised
```{r, eval=TRUE, echo=FALSE, results='hide', fig.width=2}

plot_fit_rs(df_base, column='f', title='base', ylim=c(0.5,1))

plot_fit_rs(df_ef, column='f', title='effectiveness', ylim=c(0.5,1), yblank=TRUE)

plot_fit_rs(df_ex, column='f', title='exposure', ylim=c(0.5,1), yblank=TRUE)

```

Seems that in the case where effectiveness is increased the fitnesses are closer together so does that mean that there is less selection for resistance ?

What is happening with the 2nd insecticide ?

To look at this I calculated the mean fitness of the R & S alleles for each insecticide across all genotypes. (and I repeated this both for normalised and non-normalised fitnesses). This was done by multiplying the fitness of each genotype by the number of each allele and then calculating a mean.

```{r, eval=TRUE, echo=FALSE, results='markup'}
#just to give the example of the calculation
calc_fit_rs(df_base,'f',round_places=2, verbose=TRUE)
```

The difference between the resistant and susceptible allele for each insecticide (e.g. R1-S1) was then calculated.

#Normalised
```{r, eval=TRUE, echo=FALSE}

calc_fit_rs(df_base,'f',round_places=3)
#R2-S2  R2-S2 0.066

calc_fit_rs(df_ef,'f',round_places=3)
#R2-S2  R2-S2 0.045

calc_fit_rs(df_ex,'f',round_places=3)
#R2-S2  R2-S2 0.073
```

So yes, increasing effectiveness I1 results in a decrease in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.045). This could explain the slower selection for resistance.

In contrast increasing exposure to I1 results in an increase in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.073). This could explain the faster selection for resistance.

Similar results are obtained without normalising the fitnesses first.

#not normalised
```{r, eval=TRUE, echo=FALSE}

calc_fit_rs(df_base_,'f')
#R2-S2  R2-S2 0.052

calc_fit_rs(df_ef_,'f')
#R2-S2  R2-S2 0.033

calc_fit_rs(df_ex_,'f')
#R2-S2  R2-S2 0.052

```

So in the non normalised version the difference in fitness for S&R I2 stays the same when I1 exposure is increased but decreases when I1 effectiveness is increased.

In text form :
In a mixture, increasing the exposure to one insecticide has very little effect on the selection pressure for resistance to the other. In contrast increasing the effectiveness of one insecticide in a mixture decreases the slection pressure for resistance to the other.