### 2016-11-23 Matt/AttemptPlottingRotations.Rmd

---
title: "AttemptPlottingRotations"
output: pdf_document
author: "Matt Powney"
date: "`r Sys.Date()`"
---

#The following code generates and plots the rotation resistances for Curtis figure 2.

```{r,echo=FALSE,results='hide'}
findResistancePointsRotation <- function( rotation_generations = 10,
                                          listOutI1,
                                          listOutI2,
                                          locus,
                                          criticalPoints = c(0.1,0.25,0.5)
){

  #set the number of scenarios
  numScenarios <- length(listOutI1$results)

  #create matrix to receive the results, name rows including gen_cP for generations to reach criticalPoints.
  resistGens <- matrix( nrow=length(criticalPoints), ncol=numScenarios, dimnames=list(paste0('gen_cP',criticalPoints) ) )

  #for each scenario
  for(scenarioNum in 1:numScenarios)
  {
    #for each critical point
    for(criticalPointNum in seq_along(criticalPoints) )
    {
      criticalPoint <- criticalPoints[criticalPointNum]
      #cat("criticalPoint ", criticalPoint,"\n")

      #getting results matrix out of the list for the separate insecticides
      resultsI1 <- listOutI1$results[[scenarioNum]]
      resultsI2 <- listOutI2$results[[scenarioNum]]

      #extracting the resistances for insecticides 1 and 2
      resistancesI1 <-    rowMeans(cbind(resultsI1[,'m.R1'],
                                         resultsI1[,'f.R1'] ))
      resistancesI2 <-    rowMeans(cbind(resultsI2[,'m.R2'],
                                         resultsI2[,'f.R2'] ))

      #Isolating generations where resistance hasn't been met for I1 and I2.
      gensI1 <- which( resistancesI1 < criticalPoint)
      gensI2 <- which( resistancesI2 < criticalPoint)

      #Connecting first detectable resistance generations.
      gensI1=c(gensI1, max(gensI1)+1)
      gensI2=c(gensI2, max(gensI2)+1)

      #Calculate total number of generations until resistance for rotation.
      overallgens = length(gensI1) + length(gensI2)

      #The following code calculates the final number of generations before resistance for each insecticide individually.

      roundsI1=(length(gensI1)%/%rotation_generations)
      roundsI2=(length(gensI2)%/%rotation_generations)

      leftover=(length(gensI2)%%rotation_generations)

      if (roundsI1 > roundsI2){

        I1gen=overallgens

        if (leftover==0){

          I2gen=(2*roundsI2)*rotation_generations + leftover
        }

        if (leftover!=0){

          I2gen=(2*roundsI2 + 1)*rotation_generations + leftover
        }

      }

      else 
      {
        I2gen=overallgens

        I1gen=(2*(roundsI1)*rotation_generations + leftover)

      }

      #Calculate total number of generations until resistance for rotation.
      if (locus == "both")
      {
        if (I1gen>I2gen){
          resistGens[criticalPointNum,scenarioNum] <- I1gen
        }

        if (I1gen<I2gen){
          resistGens[criticalPointNum,scenarioNum] <- I2gen
        }

      }

      #Report number of generations for insecticide 1 and insecticide 2.
      else if (locus ==1)
      {
        resistGens[criticalPointNum,scenarioNum] <- I1gen

      }

      else if (locus == 2)
      {
        resistGens[criticalPointNum,scenarioNum] <- I2gen

      }

    }
  }

  #replace any NAs with 999 to show that resistance not reached
  resistGens[is.na(resistGens)] <- 999

  return(resistGens)

}

```

```{r,echo=FALSE,results='hide'}
#To mirror the inputs for Curtis Figure 2.

exp1 = NULL
exp2 = NULL
max_gen = 500
P_1 = 0.01
P_2 = 0.01
recomb_rate = 0.5
exposure = 0.9
phi.SS1_a0 = 0
phi.SS1_A0 = 0.73
phi.SS2_0b = 0
phi.SS2_0B = 1
W.SS1_00 = 1 
W.SS2_00 = 1
h.RS1_00 = 0
h.RS1_a0 = 0
h.RS1_A0 = 0.17
h.RS2_00 = 0
h.RS2_0b = 0
h.RS2_0B = 0.0016
s.RR1_a0 = 0 
s.RR1_A0 = NULL
s.RR2_0b = 0
s.RR2_0B = NULL
rr_restoration_ins1 = 0.23/0.73
rr_restoration_ins2 = 0.43/1
z.RR1_00 = 0
z.RR2_00 = 0
sexLinked = 0
male_exposure_prop = 1
correct_mix_deploy = 1

#Set exposure and inputs for insecticide 1
a <- setExposure(exposure = exposure, exp1 = exp1, exp2 = exp2, 
                 insecticideUsed = "insecticide1", male_exposure_prop = male_exposure_prop, 
                 correct_mix_deploy = correct_mix_deploy)

inputI1 <- setInputOneScenario(a = a, max_gen = max_gen, 
                               P_1 = P_1, P_2 = P_2, recomb_rate = recomb_rate, phi.SS1_a0 = phi.SS1_a0, 
                               phi.SS1_A0 = phi.SS1_A0, phi.SS2_0b = phi.SS2_0b, phi.SS2_0B = phi.SS2_0B, 
                               W.SS1_00 = W.SS1_00, W.SS2_00 = W.SS2_00, h.RS1_00 = h.RS1_00, 
                               h.RS1_a0 = h.RS1_a0, h.RS1_A0 = h.RS1_A0, h.RS2_00 = h.RS2_00, 
                               h.RS2_0b = h.RS2_0b, h.RS2_0B = h.RS2_0B, s.RR1_a0 = s.RR1_a0, 
                               s.RR1_A0 = s.RR1_A0, s.RR2_0b = s.RR2_0b, s.RR2_0B = s.RR2_0B, 
                               rr_restoration_ins1 = rr_restoration_ins1, rr_restoration_ins2 = rr_restoration_ins2, 
                               z.RR1_00 = z.RR1_00, z.RR2_00 = z.RR2_00, sexLinked = sexLinked, 
                               male_exposure_prop = male_exposure_prop, correct_mix_deploy = correct_mix_deploy)

#Set exposure and inputs for insecticide 1

a <- setExposure(exposure = exposure, exp1 = exp1, exp2 = exp2, 
                 insecticideUsed = "insecticide2", male_exposure_prop = male_exposure_prop, 
                 correct_mix_deploy = correct_mix_deploy)

inputI2 <- setInputOneScenario(a = a, max_gen = max_gen, 
                               P_1 = P_1, P_2 = P_2, recomb_rate = recomb_rate, phi.SS1_a0 = phi.SS1_a0, 
                               phi.SS1_A0 = phi.SS1_A0, phi.SS2_0b = phi.SS2_0b, phi.SS2_0B = phi.SS2_0B, 
                               W.SS1_00 = W.SS1_00, W.SS2_00 = W.SS2_00, h.RS1_00 = h.RS1_00, 
                               h.RS1_a0 = h.RS1_a0, h.RS1_A0 = h.RS1_A0, h.RS2_00 = h.RS2_00, 
                               h.RS2_0b = h.RS2_0b, h.RS2_0B = h.RS2_0B, s.RR1_a0 = s.RR1_a0, 
                               s.RR1_A0 = s.RR1_A0, s.RR2_0b = s.RR2_0b, s.RR2_0B = s.RR2_0B, 
                               rr_restoration_ins1 = rr_restoration_ins1, rr_restoration_ins2 = rr_restoration_ins2, 
                               z.RR1_00 = z.RR1_00, z.RR2_00 = z.RR2_00, sexLinked = sexLinked, 
                               male_exposure_prop = male_exposure_prop, correct_mix_deploy = correct_mix_deploy)

```

```{r,echo=FALSE,results='hide'}
input <- cbind(inputI1, inputI2)

#Run the code for sequential inputs as the population dynamics are required to calculate the rotation resistance levels.
listOut <- runModel2(input = input, produce.plots = FALSE)

#Set the critical point 0.5.
criticalPoint=0.5

#Come back to these for test runs
#listOutCurtI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment='curtis')
#listOutCurtI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment='curtis')

#listOutCurtI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment='curtis')
#listOutCurtI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment='curtis')
#resultsI1 <- listOutCurtI1$results[[1]]
#resultsI2 <- listOutCurtI2$results[[1]]

#Extract the results for insecticides 1 and 2.
resultsI1 <- listOut$results[[1]]
resultsI2 <- listOut$results[[2]]

#extracting the resistances for insecticides 1 and 2
resistancesI1 <-    rowMeans(cbind(resultsI1[,'m.R1'],
                                   resultsI1[,'f.R1'] ))
resistancesI2 <-    rowMeans(cbind(resultsI2[,'m.R2'],
                                   resultsI2[,'f.R2'] ))

#Isolating generations where resistance hasn't been met for I1 and I2.
gensI1 <- which( resistancesI1 < criticalPoint)
gensI2 <- which( resistancesI2 < criticalPoint)

gensI1=c(gensI1,max(gensI1)+1)
gensI2=c(gensI2,max(gensI2)+1)

ins1=resistancesI1[gensI1]
ins2=resistancesI2[gensI2]

ins1
ins2

#Come back to the following to generate a function

#m=2

#rotation_generations=5

#for(i in 1:m){

#for (k in 1:rotation_generations)

#{
#fl[(2*(i-1)*rotation_generations)+k]=ins1[k]

#fl[rotation_generations*i+k]=ins1[rotation_generations*i]

#}
#}

#fl=rep(0,)

gens=seq(1,65, by=1)

i1fiverot=c(ins1[1:5],rep(ins1[6],5),ins1[7:11],rep(ins1[11],5),ins1[12:16],rep(ins1[16],5),ins1[17:21],rep(ins1[21],5),ins1[22:26],rep(ins1[26],5),ins1[27:31],rep(ins1[31],5),ins1[32:36])

i2fiverot=c(rep(ins2[1],5),ins2[2:6],rep(ins2[6],5),ins2[7:11],rep(ins2[11],5),ins2[12:16],rep(ins2[16],5),ins2[17:21],rep(ins2[21],5),ins2[22:26],rep(ins2[26],5),ins2[27:30],rep(ins2[30],6))

i1tenrot=c(ins1[1:10],rep(ins1[11],10),ins1[12:21],rep(ins1[21],10),ins1[22:31],rep(ins1[31],10),ins1[32:36])

i2tenrot=c(rep(ins2[1],10),ins2[2:11],rep(ins2[11],10),ins2[12:21],rep(ins2[21],10),ins2[22:30],rep(ins2[30],6))

i1twentyrot=c(ins1[1:20],rep(ins1[21],20),ins1[22:36],rep(ins1[36],10))

i2twentyrot=c(rep(ins2[1],20),ins2[2:21],rep(ins2[21],16),ins2[22:30])

```

## The following code makes a plot of the Curtis resistances along with rotations.

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8}

#Use the run Curtis function to generate the plots for mixtures and sequential strategies.

runcurtis_f2()

#The following code crudely generates the resistance levels for rotations required for the ASTMH presentation dependent on the resistance_generations. IMPROVE THIS.

#For 5 generation rotations
lines( gens, log10(i1fiverot*100), col=adjustcolor("red"), lty=4,lwd=3 )
lines( gens, log10(i2fiverot*100), col=adjustcolor("blue"), lty=4,lwd=3 )

#For 10 generation rotations
lines( gens, log10(i1tenrot*100), col=adjustcolor("red"), lty=4, lwd=3 )
lines( gens, log10(i2tenrot*100), col=adjustcolor("blue"), lty=4, lwd=3 )

#For 20 generation rotations
lines( gens, log10(i1twentyrot*100), col=adjustcolor("red"), lty=4, lwd=3 )
lines( gens, log10(i2twentyrot*100), col=adjustcolor("blue"), lty=4, lwd=3 )

```