### 2017-12-13 inst/documents/linkage_dis_2017.Rmd
quick look at linkage

---
  title: "linkage disequilibrium"
  output: pdf_document
  #output: word_document
  author: "Andy South"
  date: "`r Sys.Date()`"
---

### documenting whats happening with linkage disequilibrium  after review of paper2

```{r,echo=FALSE,results='hide',warning=FALSE}
  library(resistance)
  library(ggplot2)
```

### fig 8D in paper2 with a new & old insecticide, mix slower

```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left'}

listOut <- runcurtis_f2( recomb_rate=0.5, max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, maxX = 150, labelMixSeqRatio = 1 )
```

### now plot the LD for that scenario
It's not obvious to me what this shows. 
Possible that only the mixture one is meaningfull. In the sequential cases resistance to the other insecticide probably remains at 0 and therefore LD is not meaningfull.

```{r, eval=TRUE, echo=FALSE, fig.height=8, fig.width=7, fig.show='hold', fig.align='left'}

#we are mostly interested in what happens in the first 80 generations, by that time the frequency of I1 has reached 0.5 both alone and in the mixture.

#in listOut from runcurtis_f2, 1=I1, 2=I2, 3=mix
par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
par(mfrow=c(3,1))

plotlinkage(listOut$results[[3]], plot_d = TRUE, max_gen_plot = 80, main = "mixture")
plotlinkage(listOut$results[[1]], plot_d = TRUE, max_gen_plot = 80, main = "insecticide1 alone")
plotlinkage(listOut$results[[2]], plot_d = TRUE, max_gen_plot = 80, main = "insecticide2 alone")

```