### 2018-01-23 inst/documents/paper2/summary_fig_pap2.Rmd
create a labelled summary fig for paper2

---
title: "summary figure for malaria journal paper"
output: 
  pdf_document:

#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, eval=TRUE, include=FALSE}
#```{r, eval=FALSE, include=FALSE}
# options to create final publication quality figures
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='')
```

```{r,echo=FALSE,results='hide',warning=FALSE}
  library(resistance)
  library(ggplot2)
```

```{r summary_fig, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Comparing time-to-resistance for mixtures and sequences  A. Resistance arises slower for the mixture, B. Resistance arises slower for the sequence. All inputs set to 0.5 except for starting frequency of resistance set to 0.01, and effectiveness of insecticide 1 set to 0.4 in A and 0.8 in B."}

#2 panel fig A. sequence better, B. mix better

#eff1  0.4  0.8

# add side panels for labels
#layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
# 2 side by side plus lower panl for legent
#layout( matrix(c(1:2,3,3),2,2, byrow = TRUE), heights=c(1,0.04), widths=c(1,1) )
#layout.show(n=3)
# just 2 side by side plus lower panel for legent
layout( matrix(c(1:2),2,2, byrow = TRUE), heights=c(1), widths=c(1,1) )

#eff1 0.4 seq better

#plot margins, high t just a bodge to get graphs down closer to legend
par(mar=c(1.5, 4, 15, 0)) #b,l,t,r default c(5, 4, 4, 2)

runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.4 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, addStrategyLabels = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = NULL )

mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r

#plot margins, hight t just a bodge to get graphs down closer to legend
par(mar=c(1.5, 3, 15, 1)) #b,l,t,r default c(5, 4, 4, 2)

#eff1 0.8 mix better
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, addStrategyLabels = FALSE, ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = NULL )

legend( 'bottomright', legend=c("insecticide1 alone", "insecticide1 in mix", "insecticide2 alone", "insecticide2 in mix"), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.9 )

mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r

#adding legend in lower panel

# par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
# 
# plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m') 
# 
# legend( 'bottom', inset=0.1, legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   "), col=c("red","blue","red","blue"), lty=c(2,2,1,1), pch=c(NA,NA,NA,NA), bty="n", cex=1.05, horiz =TRUE )

```